---
title: "Joining Assignment"
author: "Nuria Diaz "
date: "`r format(Sys.time(), '%B %d, %Y')`"
format:
  html:
    self-contained: true
    toc: true
    toc-location: left
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)

```

# Load the datasets

```{r}

counties_demographics <- readRDS(here("data", "counties_demographics.rds"))
counties_elexresults <- readRDS(here("data", "counties_elexresults.rds"))
counties_rural_urban_suburban <- readRDS(here("data", "counties_rural_urban_suburban.rds"))



```

# Exercises

1)  

Join together the election results and urban/suburban/rural tables above, and then use the merged table to:

-   filter for just Virginia

-   group_by and summarize to calculate the mean pct of Trump 2020 vote (`pct_r_2020`) by urban/rural categories (`rural_urban`)

```{r}
inner_join(counties_demographics, counties_rural_urban_suburban) %>% 
  filter(state_code == "VA") %>% 
  group_by(rural_urban) %>% 
  summarize(average = mean(pct_r_2020))



```

2)  

Using the joined election results / urban/suburban/rural table created above:

-   filter for just Virginia

-   count up how many counties were won by party in 2020 (`winner_2020`) by urban/rural categories

```{r}


inner_join(counties_demographics, counties_rural_urban_suburban) %>% 
  filter(state_code == "VA") %>% 
  group_by(rural_urban) %>% 
  summarize(n())

newtable <- inner_join(counties_demographics, counties_rural_urban_suburban) %>% 
  filter(state_code == "VA") %>% 
  group_by(rural_urban) %>% 
  summarize(n()) 

```

3)  

Join together the election results and demographics tables above, and then use the joined table to:

-   filter for just Florida

-   group/summarize to calculate the mean pct of Trump 2020 vote in the counties above or below national average for college graduates (`pct_ed_college_all_abovebelow_natl`)

```{r}


inner_join(counties_elexresults, counties_demographics) %>% 
  filter(state_code == "FL") %>% 
  group_by(pct_ed_college_all_abovebelow_natl) %>% 

```

4)  

Using the joined table you made above:

-   group/summarize to calculate the mean pct of Trump 2020 vote by counties above or below the national average for college graduates (`pct_ed_college_all_abovebelow_natl`).

    Note we're looking for the entire country here.

```{r}

inner_join(counties_elexresults, counties_demographics) %>% 
    group_by(pct_ed_college_all_abovebelow_natl) %>% 
    summarize(average = mean(republican_2020))


```

Now using the joined datasets above, come up with two groupings you find interesting or newsworthy in a particular state, and replicate the steps above them them instead.

5)  

```{r}


#interested in seeing how nonwhite voters voted in the elections specifically the average mean for the democratic candidates in Texas. 

inner_join(counties_elexresults, counties_demographics) %>% 
  filter(state_code == "TX") %>% 
  group_by(pct_race_nonwhite) %>% 
  summarize(average = mean(democrat_2020))


```

6)  

```{r}

#looking at immigrant voting paterns in Texas specifically 
inner_join(counties_elexresults, counties_demographics) %>% 
  filter(state_code == "TX") %>% 
  group_by(pct_born_foreign) %>% 
  summarize(average = mean(democrat_2020))



```

7)  Extra credit

Up until now we've joined two tables together. But we actually have three tables in this project. How might we join all three together into a single table, so all the measures for each county are combined in one place? (Hint: you can do this is multiple steps, doesn't have to be one step only.)

```{r}



tablea <- inner_join(counties_elexresults, counties_demographics)
inner_join(tablea, counties_rural_urban_suburban)



```
